---
title: "p8105_hw3_bw2595"
author: "Blair Wigsten"
date: "10/8/2019"
output: github_document
editor_options: 
  chunk_output_type: inline
---

```{r}
library(tidyverse)
library(knitr)
```


#Problem 1
```{r}
library(p8105.datasets)
data("instacart")
instacart
```
*There are 1,384,617 observations and 15 variables in the instacart dataset. Most variables are integer variables, with the rest (eval_set, product_name, aisle, and department) as character variables. Key variables for analysis could include if the item was reordered to determine popularity, aisle name in order to filter observations based on certain criteria, and order_dow to see when most orders are placed during the week. An illustrative example of this data would be the following for the first observation: 
*illustrative examples

## How many aisles are there, and which aisles are the most items ordered from?
```{r}
instacart %>%
  count(aisle) %>%
  arrange(desc(n))
```
There are 134 distinct aisles. The aisles "fresh vegetables", "fresh fruits", and "packaged vegetables fruits" are where most items are ordered from based on the table produced above. 

## Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
items_plot = instacart %>%
  count(aisle) %>%
  filter(n > 10000) %>%
  rename(n_items_ordered = n) %>%
  ggplot(aes(x = reorder(aisle, -n_items_ordered), y = n_items_ordered)) +
    geom_bar(stat = "identity") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    labs(
    title = "Number of items ordered per aisle",
    x = "aisle name",
    y = "number of items ordered"
    )
items_plot
```
* This plot shows the number of items ordered in each aisle, limiting the plot to aisles with more than 10000 items ordered. Items have been arranged by most ordered to least ordered. 


## Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r}
popular_product = instacart %>%
  group_by(product_name, aisle) %>%
  summarize(
    n_ordered = n()) %>%
  group_by(aisle) %>%
  filter(
    aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits"), 
    min_rank(desc(n_ordered)) < 4) %>%
  arrange(n_ordered, aisle) %>%
  kable(format = "pandoc", caption = "Most popular items by aisle")
popular_product
```
* Based on the table above, snack sticks were the most ordered item within "dog food care" with 30 orders. Light brown sugar was the most ordered item in the "baking ingredients" aisle with 499 orders, and organic baby spinach was the most ordered item in the "packaged vegetables fruits" aisle with 9784 orders. 


## Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
apples_cream = instacart %>%
  rename(order_hour = order_hour_of_day) %>%
  select(product_name, order_dow, order_hour) %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(
    mean_hour_of_day = mean(order_hour)) %>%
  mutate(
    day_of_week = recode(order_dow,
                         `0` = "Sunday", 
                         `1` = "Monday",
                         `2` = "Tuesday",
                         `3` = "Wednesday",
                         `4` = "Thursday",
                         `5` = "Friday",
                         `6` = "Saturday")) %>%
  select(product_name, day_of_week, mean_hour_of_day) %>%
  pivot_wider(
    names_from = "product_name",
    values_from = "mean_hour_of_day") %>%
  kable(format = "pandoc", caption = "Mean Hour of Pink Lady Apples and Coffee Ice Cream Orders")
apples_cream
```
* The table above shows the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. 

# Problem 2

## load the brfss_smart2010 data
```{r}
library(p8105.datasets)
data("brfss_smart2010")
brfss_smart2010
```

## Data cleaning; check to see how responses are coded already
```{r}
response_table = brfss_smart2010 %>%
  filter(Topic == "Overall Health") %>%
  count(Response)
response_table
```
* responses are already in the range of "Excellent" to "Poor", taking on values of "Excellent", "Very good", "Good", "Fair", and "Poor". 

## Data cleaning 
```{r}
brfss_smart2010 = brfss_smart2010 %>%
  janitor::clean_names() %>%
  rename(
    state = locationabbr, 
    county = locationdesc,
    lcl = confidence_limit_low,
    ucl = confidence_limit_high) %>%
  separate(county, into = c("states", "county")) %>%
  select(-states, -data_value_footnote, -data_value_footnote_symbol, -location_id) %>%
  filter(topic == "Overall Health") %>%
  mutate(response = factor(response, c("Poor", "Fair", "Good", "Very good", "Excellent")))
brfss_smart2010
```
* Here, we clean the variable names, then rename locationabbr to state, confidence_limit_low to lcl, and confidence_limit_high to ucl for clarity. Then, we separate county into states and county. Then we eliminate all the repetitive variables (states) and variables with no information. Finally, we filter the data by topic "Overall Health" and mutate response into a factor variable ranging from "Poor" to "Excellent".

In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
brfss1 = brfss_smart2010 %>% 
  filter(year == "2002") %>%
  distinct(state, county) %>%
  count(state) %>%
  filter(n >= 7) %>%
  rename(number_sites = n)
brfss1

brfss2 = brfss_smart2010 %>% 
  filter(year == "2010") %>%
  distinct(state, county) %>%
  count(state) %>%
  filter(n >= 7) %>%
  rename(number_sites = n) 
brfss2
```
In 2002, there were 5 states which had 7 or more observation sites (counties), shown in the first table above. In 2010, there were 14 states which had 7 or more observations sites (counties), shown in the second table above. 

Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
excellent = brfss_smart2010 %>%
  filter(response == "Excellent") %>%
  group_by(year, state) %>%
  summarize(
    mean_data_value = mean(data_value)) %>%
  ggplot(aes(x = year, y = mean_data_value, group = state, color = state)) +
  geom_line()
excellent
```


Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State

```{r}
NY_distribution = brfss_smart2010 %>%
  filter(year %in% c("2006", "2010"), state == "NY") %>%
  select(year, response, data_value) %>%
  ggplot(aes(x = response, y = data_value)) +
    geom_boxplot() +
    facet_grid(~year) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    labs(
      title = "Distribution of Data Values for 2006 and 2010 in NY"
    )
NY_distribution
```



Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).
```{r}
accel = read_csv(file = "data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    activity_1:activity_1440, 
    names_to = "activity_number",
    values_to = "activity") %>%
  mutate(
    day_type = if_else(day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"), "day", day),
    day_type = if_else(day_type %in% c("Saturday", "Sunday"), "weekend", day_type)) %>%
  rename(
    day_number = day_id,
    day_name = day
  )
accel
```
After cleaning the names of the variables, I pivoted longer to tidy the activity variables and consolidate into activity_number and activity value as two separate variables. Then, to create the day_vs_end variable indicating weekday or weekend, I mutated the day variable to correspond to whether the value should be a weekday or weekend. The final cleaned accel dataset contains 6 variables (week, day_id, day, activity_number, activity, and day_vs_end). There are 50,400 observations in the final accel dataset.


Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r}
accel_sum = accel %>%
  group_by(day_number) %>%
  summarize(activity_sum = sum(activity))

kable(accel_sum, format = "pandoc", caption = "Total activity per day")
accel_sum

accel_sum %>%
  ggplot(aes(x = day_number, y = activity_sum)) +
  geom_line()


```
Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
plot = accel %>%
  separate(activity_number, into = c("activity_number", "minute_number")) %>%
  select(-activity_number) %>%
  group_by(minute_number = cut(minute_number, breaks = 60))
plot
```



group_by(DeviceTime = cut(DeviceTime, breaks="5 min")) %>%
  summarize(Concentration = mean(Concentration))









