---
title: "p8105_hw3_bw2595"
author: "Blair Wigsten"
date: "10/8/2019"
output: github_document
editor_options: 
  chunk_output_type: inline
---

```{r}
library(tidyverse)
library(knitr)
```

```{r}
library(p8105.datasets)
data("instacart")
instacart
```
There are 1,384,617 observations and 15 variables in the instacart dataset. Most variables are integer variables, with the rest (eval_set, product_name, aisle, and department) as character variables. 
*key variables 
*illustrative examples
```{r}
instacart %>%
  count(aisle) %>%
  arrange(desc(n))
```
There are 134 distinct aisles, with "fresh vegetables", "fresh fruits", and "packaged vegetables and fruits" are where most items are ordered from based on the table produced above. 


```{r}
items_plot = instacart %>%
  count(aisle) %>%
  filter(n > 10000) %>%
  rename(n_items_ordered = n) %>%
  ggplot(aes(x = reorder(aisle, -n_items_ordered), y = n_items_ordered)) +
    geom_bar(stat = "identity") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    labs(
    title = "Number of items ordered per aisle",
    x = "aisle name",
    y = "number of items ordered"
    )
items_plot
```


Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
popular_product = instacart %>%
  group_by(product_name, aisle) %>%
  summarize(
    n_ordered = n()) %>%
  group_by(aisle) %>%
  filter(
    aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits"), 
    min_rank(desc(n_ordered)) < 4) %>%
  arrange(n_ordered, aisle) %>%
  kable(format = "pandoc", caption = "Most popular items by aisle")
popular_product
```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
apples_cream = instacart %>%
  rename(order_hour = order_hour_of_day) %>%
  select(product_name, order_dow, order_hour) %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(
    mean_hour_of_day = mean(order_hour)) %>%
  mutate(
    day_of_week = recode(order_dow,
                         `0` = "Sunday", 
                         `1` = "Monday",
                         `2` = "Tuesday",
                         `3` = "Wednesday",
                         `4` = "Thursday",
                         `5` = "Friday",
                         `6` = "Saturday")) %>%
  select(product_name, day_of_week, mean_hour_of_day) %>%
  pivot_wider(
    names_from = "product_name",
    values_from = "mean_hour_of_day") %>%
  kable(format = "pandoc", caption = "Mean Hour of Pink Lady Apples and Coffee Ice Cream Orders")
apples_cream
```


```{r}
library(p8105.datasets)
data("brfss_smart2010")
brfss_smart2010
```

format the data to use appropriate variable names;
focus on the “Overall Health” topic
include only responses from “Excellent” to “Poor”
organize responses as a factor taking levels ordered from “Poor” to “Excellent”

```{r}
response_table = brfss_smart2010 %>%
  filter(Topic == "Overall Health") %>%
  count(Response)
response_table
```
responses are already in the range of "Excellent" to "Poor", taking on values of "Excellent", "Very good", "Good", "Fair", and "Poor". 

```{r}
brfss_smart2010 = brfss_smart2010 %>%
  janitor::clean_names() %>%
  rename(
    state = locationabbr, 
    county = locationdesc,
    lcl = confidence_limit_low,
    ucl = confidence_limit_high) %>%
  separate(county, into = c("states", "county")) %>%
  select(-states) %>%
  filter(topic == "Overall Health") %>%
  mutate(response = factor(response, c("Poor", "Fair", "Good", "Very good", "Excellent")))
brfss_smart2010
```
In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
brfss1 = brfss_smart2010 %>% 
  filter(year == "2002") %>%
  group_by(state) %>%
  filter(n() >= 7) %>%
  distinct(state, county)
brfss1 %>%
  count(state) %>%
  filter(n >= 7) %>%
  rename(number_observations = n) 

brfss2 = brfss_smart2010 %>% 
  filter(year == "2010") %>%
  separate(county, into = c("states", "county")) %>%
  group_by(state) %>%
  filter(n() >= 7) %>%
  distinct(state, county)
brfss2 %>%
  count(state) %>%
  filter(n >= 7) %>%
  rename(number_observations = n) 
```
In 2002, there were 5 states which had 7 or more observation sites (counties), shown in the first table above. In 2010, there were 14 states which had 7 or more observations sites (counties), shown in the second table above. 

Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
excellent = brfss_smart2010 %>%
  filter(response == "Excellent") %>%
  group_by(year, state) %>%
  summarize(
    mean_data_value = mean(data_value)) %>%
  ggplot(aes(x = year, y = mean_data_value, group = state)) +
  geom_line()
excellent
```


Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State

```{r}
NY_2006 = brfss_smart2010 %>%
  filter(year == "2006", state == "NY") 

NY_2010 = brfss_smart2010 %>%
  filter(year == "2010", state == "NY")

ggplot(data = NY_2006, aes(x = data_value, y = response)) +
  geom_point()
ggplot(data = NY_2010, aes(x = data_value, y = response)) +
  geom_line()
```

```{r}
brfss_smart2010 %>%
  filter(year %in% c("2006", "2010")) %>%
  select(year, response, data_value) %>%
  ggplot(aes(x = data_value, y = response)) +
  geom_point() +
  facet_grid(~year)
```


```{r}
accel = read_csv(file = "fivethirtyeight_datasets/pols-month.csv") %>%
```












